Different forms of uncertainty differently affect the evolution of social learning

### 2022-07-23 ms.tex
nearly done with revision; need N=1000 data changeover

Increased selection set size $B$ led nearly uniformly to more prevalent social
learning with all other uncertainty parameters being equal ($B$ increases along
columns in Figure~\ref{fig:mainResults}).  This is because misleading social
information is easy to overcome for two reasons: (1) larger $B$ presents a more
difficult problem for trial-and-error learning; and (2) a misidentified optimal
behavior is weighted less relative to smaller $B$, meaning agents are more likely to
explore behaviors identified as non-optimal when $B$ is larger (columns of
Figure~\ref{fig:mainResults}).  
To understand point (2), consider the case where a social learner child learns
that behavior 1 has an expected payoff of 0.8, and the rest have an expected
payoff of 0, but the optimal behavior changes between generations. Assume
the child tries behavior 1, but it pays off 0.
If $B=2$, then at within-generation $t=2$, $\ledger = (0.4, 0.0)$. The probability
the agent explores behavior 2 is now 0.02. However, if $B=10$ the agent has
$\ledger = (0.4, 0.0, \ldots, 0.0)$; the probability
the agent explores one of the nine alternative behaviors increases to 0.14.
If the agent again tries behavior 1, the probability of exploring alternatives
with $B=2$ is 0.21, but with $B=10$ the probability of exploring alternatives
greatly increases to 0.71.
% Larger $B$ caused more drift when
% $L$ was small, apparently due to the limited data available to evolution.

Payoff ambiguity had a modest effect when $\pilow = 0.1$ increased to $\pilow =
0.45$, causing social learning to be favored across more values of $u$, and across
all selection set sizes and lifespans (first two rows of
Figure~\ref{fig:mainResults}).  This indicates that increased payoff ambiguity
increases the difficulty of discerning which behavior is optimal since agents evolve
to risk social learning at greater values of $u$.  When $\pilow = 0.8$ the largest
effect is smoothing and flattening of the $\meansl$ curves, which indicates more
drift (bottom row of Figure~\ref{fig:mainResults}).  Initially this results in less
frequent evolution of social learning for the same values of $u$ and $B$. But after
the $u$ value where $\meansl \approx 0.5$, social learning evolves more frequently
than for the same $u$, $B$, and $L$. When $\pilow \to \pihigh$, agents do better and
better whether they are social learners or not. We see that this leads to less
penalty for trying social information before possibly having to correct course due
to outdated social information.

When $L$ is smaller, agents have fewer opportunities to learn through trial and
error, and so it is more difficult for them to discern which strategy, social or
asocial, is optimal. We thus saw two distinct trends emerge in the evolution of
social learning in response to changes in $L$.  First, when $\pilow$ was
sufficiently small ($\pilow=0.1, 0.45$), we observed that shorter lifespan made
social learning more beneficial (and increased $\meansl$) across some values of $u$.
When $\pilow=0.8$, shorter lifespans decreased $\meansl$ in some cases and increased
$\meansl$ in other cases. It seems that shorter lifespans in this case led to more
evolutionary drift. This can be seen in two observations. First, $\meansl$ is
reduced when $L=1$ compared to other lifespans for small values of $u$ when other
lifespans lead to $\meansl \approx 1$. At larger $u$, when $\meansl \to 0$ for other
lifespans, the $L=1$ case has non-zero $\meansl$.

@@

\mt{Need to make this section more specific in reviewing the findings.}
By systematizing four common uncertainty dimensions we developed a model of
the evolution of social learning that
predicts how different forms of uncertainty might interact to foster or suppress
social learning.  The model reproduced classic predictions that social learning can
yield higher payoffs than individual learning when uncertainty is sufficiently
limited.  By disambiguating and organizing various operationalizations of
uncertainty, we developed a more nuanced theoretical understanding of which forms
of uncertainty interact in which ways to make social learning more or less
beneficial compared to asocial learning.
Our model did not make any \emph{a priori} assumptions about whether it was
more costly for agents to be social or asocial learners, as some studies have
assumed or imposed~\cite{Rogers1988,Feldman1996,Morgan2012}. 
Instead, the uncertainty structure determined whether
or not social learning was beneficial or not, and evolution directed successive
generations towards the optimal strategy.
Such detailed theoretical models of uncertainty
are critical for understanding how humans adapt to an uncertain and
rapidly changing world, both to mitigate existential
threats~\cite{Moya2020,Jones2021}, and to capitalize on new opportunities such as
transitioning to clean energy use and
production~\cite{NatureEnergyEditorialPromisesPremises2018,Brisbois2022}.

Our model necessarily made some simplifying assumptions that, while
empirically justified in many cases, may not always hold.
\mt{Review the ones mentioned in model section: non-spatial, no within-generation
environmental variability; no horizontal learning}.
We only considered
success-biased learning, although conformity is sometimes the dominant mechanism 
for social influence of behavior well~\cite{Muthukrishna2016a,Smaldino2018b}. 
Another implicit assumption
in our model was that the number of behaviors was constant, as was the relationship
between their payoffs---one was always optimal, and the rest were equally
suboptimal. However, this fails to account for evolutionary feedback which creates
new behavioral affordances as time progresses, e.g.\ via niche 
construction~\cite{Smaldino2012,Heras-Escribano2020}. Adjusting any of these
assumptions and running new computational analyses could yield further
understanding of the evolution of social learning across diverse contexts.

>>>>

Increased selection set size $B$ led nearly uniformly to more prevalent social
learning with all other uncertainty parameters being equal ($B$ increases along
columns in Figure~\ref{fig:mainResults}).  This is because misleading social
information is easy to overcome for two reasons: (1) larger $B$ presents a more
difficult problem for trial-and-error learning; and (2) a misidentified optimal
behavior is weighted less relative to smaller $B$, meaning agents are more likely to
explore behaviors identified as non-optimal when $B$ is larger (columns of
Figure~\ref{fig:mainResults}).  
To understand point (2), consider the case where a social learner child learns
that behavior 1 has an expected payoff of 0.8, and the rest have an expected
payoff of 0, but the optimal behavior changes between generations. Assume
the child tries behavior 1, but it pays off 0.
If $B=2$, then at within-generation $t=2$, $\ledger = (0.4, 0.0)$. The probability
the agent explores behavior 2 is now 0.02. However, if $B=10$ the agent has
$\ledger = (0.4, 0.0, \ldots, 0.0)$; the probability
the agent explores one of the nine alternative behaviors increases to 0.14.
If the agent again tries behavior 1, the probability of exploring alternatives
with $B=2$ is 0.21, but with $B=10$ the probability of exploring alternatives
greatly increases to 0.71. This explanation does not apply when $L=1$, where
social learning is increasingly beneficial as $B$ increases because it is more 
valuable to be possibly misled by outdated information than it is for agents
to blindly guess which is optimal.

Payoff ambiguity had a modest effect when $\pilow = 0.1$ increased to $\pilow =
0.45$, causing social learning to be favored across more values of $u$, and across
all selection set sizes and lifespans (first two rows of
Figure~\ref{fig:mainResults}).  This is because increased payoff ambiguity increases
the difficulty, and reduces the overall benefit, of discerning which behavior is
optimal; it is worth the risk for agents to possibly learn outdated information.
When $\pilow = 0.8$ the largest effect is the smoothing and flattening of the
$\meansl$ curves, which indicates more drift (bottom row of
Figure~\ref{fig:mainResults}).  For smaller values of $u$, this results in less
frequent evolution of social learning for the same values of $u$ and $B$. But after
the value of $u$ at which $\meansl \approx 0.5$, social learning evolves more
frequently than for the same $u$, $B$, and $L$. When $\pilow \to \pihigh$, agents 
accumulate increasingly more payoffs, whether they are social learners or not. 
Greater payoff ambiguity means lower risk to try possibly outdated social
information, search for the true optimal if the learned one fails.

When $L$ is smaller, agents have fewer opportunities to learn through trial and
error, so it is more difficult for them to discern which strategy is optimal, social
or asocial. We observed two distinct trends in the evolution of social learning as
$L$ was varied.  First, when $\pilow$ was sufficiently small ($\pilow=0.1, 0.45$),
we observed that shorter lifespan selected for social learning for some values of
$u$.  When $\pilow=0.8$, shorter lifespans decreased $\meansl$ in some cases and
increased $\meansl$ in other cases due to greater evolutionary drift.  In the first
case, $\meansl$ was reduced for $L=1$ for small values of $u$, when other $L > 1$
led to $\meansl \approx 1$. At larger $u$ , the $L=1$ case yields non-zero
$\meansl$, while for the other $L > 1$, $\meansl \approx 0$.

@@

\mt{Need to make this section more specific in reviewing the findings.} By
systematizing four common uncertainty dimensions we developed a model of the
evolution of social learning that predicts how different forms of uncertainty might
interact to foster or suppress social learning.  The model reproduced classic
predictions that social learning can yield higher payoffs than individual learning
when uncertainty is sufficiently limited.  
% By disambiguating and organizing various
% operationalizations of uncertainty, we developed a more nuanced theoretical
% understanding of which forms of uncertainty interact in which ways to make social
% learning more or less beneficial compared to asocial learning. 
Detailed theoretical
models and predictions of cultural evolution under uncertainty are critical for
understanding how humans adapt to an uncertain and rapidly changing world, both to
mitigate existential threats~\cite{Moya2020,Jones2021}, and to capitalize on new
opportunities such as transitioning to clean energy use and
production~\cite{NatureEnergyEditorialPromisesPremises2018,Brisbois2022}.

% Our model did not make any \emph{a priori} assumptions about whether it was
% more costly for agents to be social or asocial learners, as some studies have
% assumed or imposed~\cite{Rogers1988,Feldman1996,Morgan2012}. 
% Instead, the uncertainty structure determined whether
% or not social learning was beneficial or not, and evolution directed successive
% generations towards the optimal strategy.

Our model necessarily made some simplifying assumptions that, while
empirically justified in many cases, may not always hold.
\mt{Review the ones mentioned in model section: non-spatial, no within-generation
environmental variability; no horizontal learning}.
We only considered
success-biased learning, although conformity is sometimes the dominant mechanism 
for social influence of behavior well~\cite{Muthukrishna2016a,Smaldino2018b}. 
Another implicit assumption
in our model was that the number of behaviors was constant, as was the relationship
between their payoffs---one was always optimal, and the rest were equally
suboptimal. However, this fails to account for evolutionary feedback which creates
new behavioral affordances as time progresses, e.g.\ via niche 
construction~\cite{Smaldino2012,Heras-Escribano2020} or cumulative cultural
evolution and dynamic network structure~\cite{Smolla2019,Derex2020}. 
Adjusting any of these assumptions and running new computational analyses could
yield further understanding of the evolution of social learning across diverse
contexts.
